---
title: "Exploratory Analysis"
output: html_notebook
---

# Load data
```{r}
# Load the dataframe
data_dir <- "/Users/robertarbon/Google Drive/Courses/Analysis-best-practice/data/original"
f_name <- paste0(data_dir, '/dili_from_anti_TB_treatment.csv')
df <- read.csv(f_name, stringsAsFactors = FALSE)
str(df)
```


# Missing values

* When you're taking the measurements yourself - not such an issue. 
* If you're collating responses from questionniars or public sources you might not understand why some values are missing. 
* There are different types of *missingness*:
  1. *Default values*: the missing values are actually default values such as zero or one. 
  2. *Not applicable*: the missing values are not actually needed. e.g. pregnancy status for people without wombs. 
  3. *Missing at random*: there is no pattern to the missing values
  5. *Not missing at random*: there is some underlying reason/pattern to the missing values. 
  
* It's usually better to imput the missing values, but must work out what type they are.

## Visualization


Now use the `visdat` library to plot dataframe

```{r}
# Load libraries
library(visdat)

# plot: 
vis_dat(df)

```

# Descriptive statistics

In order to describe our data we're going to use packages from the [`tidyverse`](https://www.tidyverse.org/).  


## Table 1

Let's try to reproduce the  

```{r}
library(tidyverse)
```

### Summarize count factors 



```{r}
# Gathers up the factors we're going to count: 
tmp1 <- gather(df, key, value, c(sex, agegroup, tb_site))

#
# tmp2 <- group_by(tmp1, pdili, key, value)
head(tmp1)
# tmp3 <- tally(tmp2)


```




