---
title: "Loading and tidying data"
output:
  pdf_document: default
  html_notebook: default
---

# Load data 
```{r}
suppressMessages(library("tidyverse"))

library(tidyverse)

```

Import necessary packages upfront. 


```{r}
data_dir <- "/Users/robertarbon/Google Drive/Courses/Analysis-best-practice/data/original"
output_dir <- "/Users/robertarbon/Google Drive/Courses/Analysis-best-practice/data/processed"
```

* Put variables up front: 
  * Easy to change if directory structure changes
  * Makes code easy to read. 

```{r}
df <- read.csv(paste(data_dir, "dili_from_anti_TB_treatment.csv", sep = "/"))
head(df, 2)
```

* Load data into data frame
* Check the `head` of the dataframe to see column titles have been read in correctly. 

# Inspect the dataframe

```{r}
str(df)
```

* `str` gives:
  * a description of the variables (`int`/`num`/`Factor` etc.)
  * the size of the dataframe 
  * the first few observations of each variable

```{r}
summary(df)
```

* `summary` give a summary of each variable according to its type: 
  * `int`/`num` - min, max, quartiles 
  * `factor` - number in each factor.  
* Also gives number of missing values (`NA`)

# Cleaning the data

There are a number of problems (of varying degrees) with the dataframe as it stands.  

2. Some of the variables are miss-typed e.g. `pdili` should be a factor, not integer. 
1. There are missing values
3. The `agegroup` category labels aren't informative. 

## Miss-typed variables

```{r}
int_to_factor <- c('pdili', 't1dili', 'hiv', 'hcv', 'hbv', 'alcohol', 'liver_disease', 'tb_site')

df[int_to_factor] <- lapply(df[int_to_factor], factor)

str(df)
```

## Missing values

* Filling-in or *imputing* missing values is difficult and context dependent.  We shall save resolving this in **Exploratory data analysis**. 

* Only remove any observations that are totally missing. 

```{r}
# Select all columns that aren't the id variable. 
variables <- ! colnames(df) %in% c("id") 

# Calculate the number of variables. 
num_variables = length(variables)

# Sum the number of missing variables in each observation
num_na = rowSums(is.na(df[,variables]))

# Remove any rows that are completely missing
print(paste("Num obs before removal: ", dim(df)[1])) 
df <- df[!num_variables == num_na, ]
print(paste("Num obs after removal : ", dim(df)[1]))
```


## Relabel factors 

* `18/29` should really be `18-29`
* `40+` should really have an upper limit (see metadata and/or paper). Authors don't supply that information, so we'll leave it for now. 

```{r}
df$agegroup <- gsub("/", "-", df$agegroup)
str(df$agegroup)
```

## Change index

```{r}
# Check there are no NAs in 'id' variable. 
summary(df$id)
# Check wether id variable is unique
print(paste("Is ID unique? ", length(df$id) == length(unique(df$id))))
row.names(df) <- df$id
df <- df[,!names(df) %in% c("id")]
tail(df,2)
```

## Summary 

This can all be accomplished in easy way using the more obscure features of `read.csv`: 

```{r}
file_name = paste(data_dir, "dili_from_anti_TB_treatment.csv", sep = "/")
# Specify 'na.strings' so that it doesn't interpret missing values as a specific factor level. 
df2 <- read.csv(file_name,
               row.names='id', 
               colClasses=c('pdili'='factor', 
                            't1dili'='factor', 
                            'hiv'='factor', 
                            'hcv'='factor', 
                            'hbv'='factor', 
                            'liver_disease'='factor', 
                            'alcohol'='factor',
                            'tb_site'='factor'), 
               na.strings=c(""))
df2$agegroup <- gsub("/", "-", df2$agegroup)
str(df2)
```

# Save date frame

```{r}
out_name = paste(output_dir, 'dili_df.Rda', sep='/')
save(df2, file=out_name)
```