---
title: "Missing-values"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load data
```{r}
# Load the dataframe
data_dir <- "/Users/robertarbon/Google Drive/Courses/Analysis-best-practice/data/original"
f_name <- paste0(data_dir, '/dili_from_anti_TB_treatment.csv')
df <- read.csv(f_name, stringsAsFactors = FALSE)

```

# Missing values

* When you're taking the measurements yourself - not such an issue. 
* If you're collating responses from questionniars or public sources you might not understand why some values are missing. 
* There are different types of *missingness*:
  1. *Default values*: the missing values are actually default values such as zero or one. 
  2. *Not applicable*: the missing values are not actually needed. e.g. pregnancy status for people without wombs. 
  3. *Missing at random*: there is no pattern to the missing values
  5. *Not missing at random*: there is some underlying reason/pattern to the missing values. 
  
* It's usually better to imput the missing values, but must work out what type they are.

## Visualization


Now use the `visdat` library to plot dataframe

```{r}
# Load libraries
library(visdat)

# plot: 
vis_dat(df)

```